# Generate DATA:

```{r}
n <- 1000 # Number of samples
k <- 4 # Number of features (Not counting the intercept)
```

```{r}
set.seed(20)
library(mvtnorm)
X <- matrix(data = NA, nrow = n, ncol = k)
X <- rmvnorm(n, mean = c(1,4,6,2))
intercept = TRUE
if (intercept){
  X <- cbind(rep(1,n), X)
  k <- k+1
}
```

The coefficients $\beta_0, \dots, \beta_k$ for the linear predictor are chosen as follows,

$$
\mu  = 3X_{intercept} + 19X_1 + 4X_2 + 9X_3 + 12X_4 \quad \text{or} \quad \mu  = 19X_1 + 4X_2 + 9X_3 + 12X_4
$$

```{r}
if (intercept) {beta <- matrix(c(3, 19, 4, 9, 12), ncol = 1)
}else {beta <- matrix(c(19, 4, 9, 12), ncol = 1)}

#beta <- matrix(c(19,4,9,12), ncol = 1)
mu <- X %*% beta
```

For the choice of $\eta_1,\eta_2 \text{ and } \phi$ we use a randomly generated sample form the Dirichlet distribution

```{r}
library(DirichletReg)
#etaphi <- sqrt(rdirichlet(1,c(.5,.5,.5)))
etaphi <- c(0.149, 0.406, sqrt(0.813))
eta1 <- etaphi[1]
eta2 <- etaphi[2]
phi <- etaphi[3]
```

The last parameters to be chosen are the mixture weight $p$ and the standard deviation $\sigma$

```{r}
p <- 0.2
sd <- 100
sigma <- matrix(data = rep(sd, n), ncol = 1)
```

Using all the previous parameters we compute the parameters of the mixture

```{r}
mu1 <- mu - sigma * phi * (sqrt(1-p)/sqrt(p))
mu2 <- mu + sigma * phi * (sqrt(p)/sqrt(1-p))
sigma1 <- sigma * (eta1/ sqrt(p))
sigma2 <- sigma * (eta2/sqrt(1-p))

source("rnormix.R")
#set.seed(20)
Y <- rnormix(n, p, mu1, sigma1, mu2, sigma2)$samples
Y_mean <- rnormix(n,p, mean(mu1), mean(sigma1), mean(mu2), mean(sigma2))$samples
```

```{r}
library(ggplot2)
ggplot()+
  geom_histogram(aes(Y), bins = 90, fill = "blue", alpha = .5) +
  geom_histogram(aes(Y_mean), bins = 90, fill = "green", alpha = .5) +
  ggtitle("Histogram of Y")


```

# Parameters

```{r}
Tau <- list("tau1"=.5, "tau2"=.5) # Mixture weights
Alpha <- c(.5,.5,.5) # eta1, eta2, phi
ksi <- list("shape"=1.5, "rate"=2) # Standard deviation
beta_MLE <- solve(t(X)%*%X)%*%t(X)%*%Y
beta_hat <- matrix(beta_MLE, ncol = 1) # Mean 
```

# Sampling

```{r}
T = 1000 # Sampling epochs
# T_adapt = 100 # Adaptation epochs
# N_chains = 30
# TempMin = 0.001
# TempMax = 1
```

```{r}
source("FactoredMixReg.R")
Rprof("myProfile.out")  # Start profiling
L = MixReg_factor(T, X, Y, Tau, Alpha, ksi, beta_hat, 1)
Rprof(NULL)             # Stop profiling
summaryRprof("myProfile.out")  # Summarize results
```

```{r}
plot(L$ac_rate_p[1,],type = "l")
hist(L$P[1,100:dim(L$P)[2]])
plot(L$ac_rate_etaPhi_sqrd[1,], type = "l")
plot(L$etaPhi[1,100:dim(L$P)[2],3], type = "l")
plot(L$ac_rate_sigma_sqrd[1,], type = "l")
plot(sqrt(L$sigma_sqrd[1,]), type = "l")
plot(L$ac_rate_beta[1,],type = "l")

for (i in 1:dim(L$Beta_chains)[3]){
  plot(L$Beta_chains[1,,i], type = "l")
}


colMeans(colMeans(L$Beta_chains))
```

```{r}
source("MixReg.R")
#Temperature = seq(TempMin,TempMax, length.out = N_chains)
Temperature <- c(1,0.5^c(seq(1,4)))
Res <- list()
for (i in 1:length(Temperature)){
  success <- FALSE  # Flag to track success
  
  while (!success) {
    print(i)
    success <- tryCatch(
      {
        Res[[i]] = MixReg_factor(T, X, Y, Tau, Alpha, ksi, beta_hat, Temperature[i])
        TRUE
      },
      error = function(e) {
        cat("An error occurred. Retrying...")
        FALSE  
      }
    )
  }
  cat("Success! The loop completed without errors.\n")
}

```

## Mixture weights P:

### Chains:

```{r}
j <- 1
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$P[i,100:dim(Res[[j]]$P)[2]], type = "l", ylim = c(0,1), xlab = "p")
```

```{r}
for (i in 1:Res[[j]]$N_chains)
  hist(Res[[j]]$samples$P[i,1000:dim(Res[[j]]$samples$P)[2]],nclass = 100, xlim = c(0,1))
```

### Acceptance rates:

```{r}
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$acceptance$P[i,1:dim(Res[[j]]$acceptance$P)[2]], type = "l", ylim = c(0,1))
Res[[j]]$eps$P
```

### Sigma:

#### Chains:

```{r}
j <- 1
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$sigma[i,100:dim(Res[[j]]$sigma)[2]], type = "l", ylab = "sigma")
```

## Phi:

### Chains:

```{r}
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$acceptance$etaPhi[i,], type = "l", ylim = c(0,1))
Res[[j]]$eps$etaphi
```

## \#

```{r}
for (i in 1:Res[[j]]$N_chains){
  plot(Res[[j]]$etaPhi[i,100:dim(Res[[j]]$etaPhi)[2],3], type = "l", ylim = c(-1,1))
  abline(h=phi, col = "red")
  abline(h=-phi, col = "red")
}
```

```{r}
# for (i in 1:Res$N_chains){
#   hist(Res$samples$etaPhi[i,1000:dim(Res$samples$etaPhi)[2],component], nclass = 500, xlim = c(-1,1))
#   if (component != 3){
#       abline(v=etaphi[1], col = "red")
#       abline(v=etaphi[2], col = "red")
#   }
#   else{
#       abline(v=etaphi[component], col = "red")
#       abline(v=-etaphi[component], col = "red")
#   }
# }
# Res$eps$etaphi
component <- 2
for (i in 1:Res[[j]]$N_chains)
  hist(Res[[j]]$etaPhi[i,100:dim(Res[[j]]$etaPhi)[2],component],nclass = 100, xlim = c(0,1))
```

```{r}
j <- 1
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$etaPhi[i,100:dim(Res[[j]]$etaPhi)[2],component], type = "l", ylim = c(0,1), xlab = "p")
```

## Beta

#### Acceptance

```{r}
for (i in 1:Res[[j]]$N_chains)
  plot(Res[[j]]$acceptance$beta[i,], type = "l", ylim = c(0,1))
Res[[j]]$eps$beta
```

```{r}
#View(Res$samples$beta[,1000,])
A <- apply(Res[[j]]$beta[,400:dim(Res[[j]]$beta)[2],], 3, FUN = rowMeans)

weight = 1
for (i in 1:Res[[j]]$N_chains){
  plot(Res[[j]]$samples$beta[i,1000:dim(Res[[j]]$samples$beta)[2],weight], type = "l")
  abline(h=beta[weight], col = "red")
}
beta
beta_MLE
Res$eps$beta
```

```{r}
library(coda)
gelman <- numeric(length = length(Res))
L <- list()
for(j in 1:(length(Res)-1)){
  chain_list <- list()
  for (i in 1:Res[[j]]$N_chains){
    chain <- mcmc(cbind(as.matrix(Res[[j]]$samples$P[i,]), Res[[j]]$samples$etaPhi[i,,], as.matrix(Res[[j]]$samples$sigma[i,]), Res[[j]]$samples$beta[i,,]))
    chain_list[[i]] <- chain
  }
  L[[j]] <- mcmc.list(chain_list)
  gelman[j] <- gelman.diag(L[[j]])$mpsrf
}
gelman
traceplot(L[[j]][2])
```

```{r}
gelman.plot(L[[j]], autoburnin = TRUE )

```

```{r}
gelman.diag(L[[j]])
```

```{r}
chain <- L[[j]][[2]]
beta_pred <- matrix(colMeans(apply(Res[[j]]$samples$beta[,400:dim(Res[[j]]$samples$beta)[2],], 3, FUN = rowMeans)), ncol = 1)
mu1 <-  X %*% beta_pred - matrix(chain[,5], ncol = 1) * matrix(chain[,4], ncol = 1) * (sqrt(1-matrix(chain[,1], ncol = 1))/sqrt(matrix(chain[,1], ncol = 1)))
mu2 <- X %*% beta_pred + matrix(chain[,5] * matrix(chain[,4] * (sqrt(matrix(chain[,1])/sqrt(1-matrix(chain[,1]))
sigma1 <- matrix(chain[,4] * (matrix(chain[,2]/ sqrt(matrix(chain[,1]))
sigma2 <- matrix(chain[,4] * (matrix(chain[,3]/sqrt(1-matrix(chain[,1]))
```
